---
title: "CfDRA"
author: "Ar-Rafi Khan"
date: "`r paste('First developed on 15 March 2025. Updated on', format(Sys.Date(), '%d %B %Y'))`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Group Work Task: Data Requirements & Data Dictionary

### Task 1: Public Health in Bangladesh – Dengue Outbreak Prevention

#### Research Objective
> *"To predict dengue outbreaks in Bangladesh as a function of climate, mosquito population, and healthcare variables so that we can recommend a control measure to limit the future dengue."*

#### Data Dictionary
Before taking decision which variable to incorporate, an extensive literature search is required involving dengue specialist. Below are few results - 

* Prediction of dengue annual incidence using seasonal climate variability in Bangladesh between 2000 and 2018 (https://doi.org/10.1371/journal.pgph.0000047)

<br>

<br>

```{r, echo = FALSE, results = 'asis'}
dd1 <- matrix(NA, nrow = 4, ncol = 5)

#####Column name
colnames(dd1) <- c("Variable Name", "Description", "Data Type", "Possible Values/Format", "Data Source")

#####Populating column
dd1[1, ] <- c("Confirmed dengue cases", "Monthly confirmed dengue cases ", "Continuous variable", "Count data", "Directorate General of Health Services (DGHS)")

dd1[2, ] <- c("Monthly minimum temperature", "Temperature measured in degrees Celsius", "Continuous variable", "Range from 7 to 40 degrees Celsius", "Bangladesh Meteorological Department (BMD)")

dd1[3, ] <- c("Rainfall", "Rainfall measured in millimeters", "Continuous variable", "Numeric data", "Bangladesh Meteorological Department (BMD)")

dd1[4, ] <- c("Sunshine duration", "Sunshine duration measured in hours", "Continuous variable", "Numeric data", "Bangladesh Meteorological Department (BMD)")


#####Print the table
library(knitr)
kable(dd1,  align = c('l', 'l', 'l', 'l', 'l'),
      caption = "_Data dictionary_") 
```


Similarly we can look into other manuscript or report to get idea what and which type data should be collected before we proceed for the study.  

<br>

<br>

Some other reference - 
* Dengue and severe dengue - https://www.who.int/news-room/fact-sheets/detail/dengue-and-severe-dengue  
* Dengue Outbreak Prediction in Bangladesh based on Meteorological Factors - https://papers.ssrn.com/sol3/papers.cfm?abstract_id=5078257  




<br>



<br>

## Group Work Task: Data Processing 

### Data - VEGIEHAT Pilot Database


```{r, message = FALSE}

#####remove all list
rm(list=ls())

##### Load library
library(tidyverse)
library(readxl)

```


Load dataset from github repository


```{r, message = FALSE}
####Path of the file
url <- "https://raw.githubusercontent.com/jaynal83/dsjr/main/data/VEGIEHAT-Pilot-Database.xlsx"

#####Download the file temporarily and read it
temp_file <- tempfile(fileext = ".xlsx")
download.file(url, temp_file, mode = "wb")
vegiehat_raw <- read_excel(temp_file)

vegiehat_raw <- vegiehat_raw %>%
  select("SubmissionId", "SubmissionTime", "UserId", "DistrictName", "UpazilaName", "ItemsToChoose", "Value - Rice", "Value - Flour", "Value - Lentil", "Value - Soybean Oil", "Value - Salt", "Value - Sugar", "Value - Eggs", "Value - Chicken", "Value - Potato", "Value - Eggplant", "Value - Onion", "Value - Green Chilli", "Your Comments" ) %>%
  rename(Comments = "Your Comments")



# View the data
str(vegiehat_raw)

```

Generating time and date variable

```{r, message = FALSE}

vegiehat_raw <- vegiehat_raw %>%
  mutate(SubmissionDateOnly = as.Date(SubmissionTime), 
         SubmissionTimeOnly = format(SubmissionTime, "%H:%M")) %>%
  select(-SubmissionTime, -ItemsToChoose)

head(vegiehat_raw, 10)
```


Reshape data to long format

```{r, message = FALSE}

vegiehat <- vegiehat_raw %>%
  pivot_longer(
    cols = starts_with("Value"), 
    names_to = "Item",
    values_to = "Price"
  )

head(vegiehat_raw, 10)

```

Separate the items into two separate columns

```{r, message = FALSE}

vegiehat <- vegiehat %>%
  separate(Item, into = c("ItemsToChoose1", "ItemsToChoose2"), sep = " - ")

```

Dropping item that has no price

```{r, message = FALSE}

vegiehat <- vegiehat %>%
  rename(ItemsToChoose = ItemsToChoose2) %>%
  filter(!is.na(Price)) 
                               
head(vegiehat_raw, 10)
```

Name of the item to assign unit

```{r, message = FALSE}

table(vegiehat$ItemsToChoose)

```

Unit of the purchase option (Either 1 kg / 1L as appropriate)

```{r, message = FALSE}

vegiehat <- vegiehat %>%
    mutate(PurchaseUnit = ifelse(ItemsToChoose == "Eggs", "4 piece", 
                                 ifelse(ItemsToChoose == "Soybean Oil", "1L", "1kg")
                                 ) )

```


Keeping only requested variables

```{r, message = FALSE}

vegiehat <- vegiehat %>%
  select(SubmissionId, SubmissionDateOnly, SubmissionTimeOnly, UserId, DistrictName, UpazilaName, ItemsToChoose, PurchaseUnit, Price, Comments)

```


Converting DistrictName and UpazilaName to factor variable

```{r, message = FALSE}

vegiehat <- vegiehat %>%
  mutate(DistrictName = as.factor(DistrictName),
         UpazilaName = as.factor(UpazilaName)
         )

```

Keeping only required data file

```{r, message = FALSE}

rm(list = setdiff(ls(), "vegiehat"))

head(vegiehat, 15)
```

Structure of the dataset

```{r, message = FALSE}

str(vegiehat)

```




