---
title: "CfDRA Day1 Task1"
author: "Ar-Rafi Khan"
date: "`r paste('First developed on 15 March 2025. Updated on', format(Sys.Date(), '%d %B %Y'))`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Group Work Task: Data Requirements & Data Dictionary

### Task 1: Public Health in Bangladesh – Dengue Outbreak Prevention

#### Research Objective
> *"To predict dengue outbreaks in Bangladesh as a function of climate, mosquito population, and healthcare variables so that we can recommend a control measure to limit the future dengue."*

#### Data Dictionary
Before taking decision which variable to incorporate, an extensive literature search is required involving dengue specialist. Below are few results - 

* Prediction of dengue annual incidence using seasonal climate variability in Bangladesh between 2000 and 2018 (https://doi.org/10.1371/journal.pgph.0000047)

<br>

<br>

```{r, echo = FALSE, results = 'asis'}
dd1 <- matrix(NA, nrow = 4, ncol = 5)

#####Column name
colnames(dd1) <- c("Variable Name", "Description", "Data Type", "Possible Values/Format", "Data Source")

#####Populating column
dd1[1, ] <- c("Confirmed dengue cases", "Monthly confirmed dengue cases ", "Continuous variable", "Count data", "Directorate General of Health Services (DGHS)")

dd1[2, ] <- c("Monthly minimum temperature", "Temperature measured in degrees Celsius", "Continuous variable", "Range from 7 to 40 degrees Celsius", "Bangladesh Meteorological Department (BMD)")

dd1[3, ] <- c("Rainfall", "Rainfall measured in millimeters", "Continuous variable", "Numeric data", "Bangladesh Meteorological Department (BMD)")

dd1[4, ] <- c("Sunshine duration", "Sunshine duration measured in hours", "Continuous variable", "Numeric data", "Bangladesh Meteorological Department (BMD)")


#####Print the table
library(knitr)
kable(dd1,  align = c('l', 'l', 'l', 'l', 'l'),
      caption = "_Data dictionary_") 
```


Similarly we can look into other manuscript or report to get idea what and which type data should be collected before we proceed for the study.  

<br>

<br>

Some other reference -  

* Dengue and severe dengue - https://www.who.int/news-room/fact-sheets/detail/dengue-and-severe-dengue  
* Dengue Outbreak Prediction in Bangladesh based on Meteorological Factors - https://papers.ssrn.com/sol3/papers.cfm?abstract_id=5078257  



